Bow shocks, bow waves, and dust waves. IV. Bow shape statistics

### 2019-04-20 papers/Papers1234/obs-bowshocks.tex
Finish going through sec 2.1 and 2.2 of Paper IV

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney et al.]{
  William J. Henney,\thanks{w.henney@irya.unam.mx}
  Jorge A. Tarango-Yong,
  Luis \'Angel Guti\'errez-Soto,
  \& S. J. Arthur
  \\
  \AddressCRyA
}

>>>>

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney et al.]{
  William J. Henney,\thanks{Email: w.henney@irya.unam.mx}
  Jorge A. Tarango-Yong,
  Luis \'Angel Guti\'errez-Soto,\thanks{%
    Current address: Observatório do Valongo,
    Universidade Federal do Rio de Janeiro,
    Ladeira Pedro Antonio 43, 20080-090 Rio de Janeiro, Brazil}
  \& S. J. Arthur
  \\
  \AddressCRyA
}

### 2019-04-20 papers/Papers1234/sec-quadrics-observations.tex
Finish intro for Paper IV
Finish going through sec 2.1 and 2.2 of Paper IV

We developed a methodology for automatically tracing the arcs as follows:

@@

  of the image.

@@

  \(C_1\).  The results are shown as red and blue dots superimposed
  on the images in Figure~\ref{fig:mipsgal-examples}. Each of the two

@@

  \(C_{k,\text{mean}}\) points into a single set, \(C_{k}\), for
  the following steps.

@@

  \(R_k\) for \(|\theta_k| < 45\degr\).

@@

@@

\subsection{Subjective evaluation of the fit quality}
\label{sec:subj-eval-fit}

After these automatic steps, we subjectively evaluate the results by giving a star rating to each source:

@@

\paragraph*{5 stars} The highest-quality fit, usually corresponding to
large, sharply defined bow shocks, whose shape is determined with high
precision. An example of a 5-star fit is shown in
Figure~\ref{fig:mipsgal-examples}c.

\bigskip
%
Figure~\ref{fig:mipsgal-examples}d compares the bow shock size,
\(R_0\), determined by our fits (vertical axis) with the corresponding
value given in the \citet{Kobulnicky:2016a} catalog (horizontal axis).
For most sources with 3-star or higher rating, the two estimates agree
to within \(\pm 20\%\), but there are a small number of sources with a
discrepancy of more than a factor of two.  In all cases that we
checked, we believe that our estimates of \(R_0\) are more accurate
than those in the catalog.  It is apparent that the star ratings are
correlated with the bow shock size, with larger bow shocks tending to
receive higher ratings, although there is considerable overlap.  In
particular, most of the 1- and 2-star sources are close to the
resolution limit of the MIPSGAL \SI{24}{\um} images (\(6\arcsec\),
indicated by the dotted horizontal line in te figure).

@@

The derived bowshock shapes of all the 3-, 4-, and 5-star sources are
shown in Figure~\ref{fig:mipsgal-shapes} on the plane of \(R_c/R_0\)
versus \(R_{90} / R_0\).  Panel a shows the individual points
superimposed on the theoretical results for quadrics of revolution
(see figure caption and \S~\ref{sec:conic}), while panel b shows
contours of the kernel density estimator (KDE, see
\citealp{Leiva-Murillo:2012a, Scott:2015a}) of the two-dimensional
distribution of points on the \(R_c\)--\(R_{90}\) plane.  The
horizontal axis corresponds to the shape of the head of the bow shock
near its apex, ranging from sharper, pointier shapes with
\(R_c/R_0 < 1\) to flatter, snubber shapes with \(R_c/R_0 \gg 1\), where
it must be understood that all judgments of sharpness/flatness are
with respect to the axial separation, \(R_0\), between the source and
the bow shock apex.  The vertical axis corresponds to the shape of the
bow shock wings, ranging from closed ``C'' shapes for smaller values
of \(R_{90} / R_0\) to open ``V'' shapes for larger values of
\(R_{90} / R_0\).  The boundary between closed and open corresponds to
the paraboloids, and is shown by the solid curved line that divides
the shaded from the unshaded regions of the graph.

The KDE contours in Figure~\ref{fig:mipsgal-shapes}b show that the
distribution of 3-star sources is very similar to that of 4- and
5-star sources, although the higher-rated sources are shifted slightly
to the upper right.  Possible reasons for this are discussed in
\S~\ref{sec:corr-shape} below.
% This is probably due to higher-rated sources
% being on average bigger, as will be discussed further below.
The bulk of the sources are concentrated around the paraboloid line,
with \(1 < R_c/R_0 < 2.5\), and \(1.2 < R_{90}/R_0 < 2\).  But
significant minorities are found in three other regions: (1)~a clump
with \(R_c/R_0 \la 1\); (2)~a vertical spur towards higher \(R_{90}\)
at \(R_c/R_0 \approx 2\); and (3)~a broad horizontal tail towards higher
\(R_c\) at \(R_{90}/R_0 \la 2\).

\subsection{Correlation between bow shock size and other parameters}
\label{sec:corr-size}

>>>>

We have developed a method for automatically tracing the arcs and
determining their planitude and alatude, which is an extension of the
method described in Appendix~E of Paper~0.  The steps of the method
are as follows:

@@

  of the image. In our implementation of the method we use functions
  from the Python library \texttt{astropy.coordinates}
  \citep{Astropy-Collaboration:2018a}.

@@

  \(C_1\).  The results are shown as red and blue dots superimposed on
  the images in Figure~\ref{fig:mipsgal-examples}. Each of the two

@@

  \(C_{k,\text{mean}}\) points into a single set, \(C_{k}\), for the
  following steps.  We have found that the use of the step-back point
  greatly improves the reliability of the method, as compared with the
  simpler option of tracing rays from the stellar source.

@@

  \(R_k\) for \(|\theta_k| < 45\degr\).\label{step:Rc}

@@

\item The planitude is determined as \(\Pi = R_c / R_0\) and the mean
  alatude as \(\Lambda = \frac12 (R_{90+} + R_{90-}) / R_0\).  We also
  calculate an alatude asymmetry as
  \(\Delta\Lambda = \frac12 (R_{90+} - R_{90-}) / R_0\). Note that in this paper
  we simplify the notation of Paper~0 by no longer using the prime
  symbol (\('\)) to distinguish projected from intrinsic quantities.
  All planitudes and alatudes should be understood as specifying the
  projected bow shock shape on the plane of the sky unless noted
  otherwise.\label{step:Pi-Lambda}

@@

% \subsubsection{Subjective evaluation of the fit quality}
\label{sec:subj-eval-fit}
After these automatic steps, we subjectively evaluate the resulting
fit quality by giving a star rating to each source:

@@

\paragraph*{5 stars} The highest-quality fit, usually corresponding to
large, sharply defined bow shocks, whose shape is determined with high
precision. An example of a 5-star fit is shown in
Figure~\ref{fig:mipsgal-examples}c.
\bigskip

% 
Figure~\ref{fig:mipsgal-examples}d compares the bow shock size,
\(R_0\), determined by our fits (vertical axis) with the corresponding
value given in the \citet{Kobulnicky:2016a} catalog (horizontal axis).
For most sources with 3-star or higher rating, the two estimates agree
to within \(\pm 20\%\), but there are a small number of sources with a
discrepancy of more than a factor of two.  In all cases that we
checked, we believe that our estimates of \(R_0\) are more accurate
than those in the catalog.  It is apparent that the star ratings are
correlated with the bow shock size, with larger bow shocks tending to
receive higher ratings, although there is considerable overlap.  In
particular, most of the 1- and 2-star sources are close to the
resolution limit of the MIPSGAL \SI{24}{\um} images (\(6\arcsec\),
indicated by the dotted horizontal line in te figure).

@@

The derived bowshock shapes of all the 3-, 4-, and 5-star sources are
shown in Figure~\ref{fig:mipsgal-shapes} on the \(\Pi\)--\(\Lambda\) plane,
which was discussed extensively in Paper~0.  Panel a shows each
individual source with horizontal error bars that represent the
dispersion of traced points from the circle fit (step~\ref{step:Rc} of
the previous section) and vertical error bars that show the wing
asymmetry \(\Delta\Lambda\) (step~\ref{step:Pi-Lambda}).  The horizontal error
bars, which are a proxy for observational uncertainties, become very
small for the 4- and 5-star sources, whereas the vertical error bars
remain at roughly the \(10\%\) level since they reflect real
asymmetries in the shapes of the bow shocks.

The horizontal axis corresponds to the shape of the head of the bow
shock near its apex, ranging from sharper, pointier shapes with
\(\Pi < 1\) to flatter, snubber shapes with \(\Pi \gg 1\), where it must be
understood that all judgments of sharpness/flatness are with respect
to the axial separation, \(R_0\), between the source and the bow shock
apex.  The vertical axis corresponds to the shape of the bow shock
wings, ranging from closed ``C'' shapes for smaller values of
\(\Lambda\) to open ``V'' shapes for larger values of \(\Lambda\).  The background
shading shows the regions of the \(\Pi\)--\(\Lambda\) plane occupied by simple
quadric shapes (see \S~4 of Paper~0): oblate spheroids (light gray),
prolate spheroids (dark gray), and hyperboloids (white). The boundary
between closed and open corresponds to the paraboloids, and is shown
by the solid curved line that divides the dark gray and white regions
of the graph.

In Figure~\ref{fig:mipsgal-shapes}b, we present a smoothed version of
the same data.  The contours show the kernel density estimator (KDE,
see \citealp{Leiva-Murillo:2012a, Scott:2015a}) of the two-dimensional
distribution of points on the \(\Pi\)--\(\Lambda\) plane.  The KDE contours
indicate that the distribution of 3-star sources is very similar to
that of 4- and 5-star sources, although the higher-rated sources are
shifted slightly to the upper right.  Possible reasons for this are
discussed in \S~\ref{sec:corr-shape} below.
% This is probably due to higher-rated sources
% being on average bigger, as will be discussed further below.
The bulk of the sources are concentrated around the paraboloid line,
with \(1 < \Pi < 3\), and \(1.2 < \Lambda < 2\).  But significant minorities
are found in three other regions: (1)~a clump with \(\Pi \la 1\); (2)~a
vertical spur towards higher \(\Lambda\) at \(\Pi \approx 1.8\); and (3)~a broad
horizontal tail towards higher \(\Pi\) at \(\Lambda \la 2\).  Thumbnail images
of the \SI{24}{\um} emission of selected sources from different
regions of the distribution are shown, in order to illustrate the
range of bow shock morphologies encountered.

\subsection{Correlations between non-shape parameters of the OB sources}
\label{sec:corr-size}
